---
title: "COMLEX2_Advising_Report_Template"
author: "Sam Manski"
format: 
  pdf:
    fontsize: 9
---

{{< pagebreak >}}

```{r}
#| echo: false
#| results: "asis"
#| include: !expr (!advising)

cat("## ")
cat(data$First_Name, " ", data$Last_Name)
cat(" COMLEX-USA Level 2-CE Prediction Report")
cat("\n")
```

```{r}
#| echo: false
#| results: "asis"
#| include: !expr advising

cat("## ")
cat(data$First_Name, " ", data$Last_Name)
cat(" COMLEX-USA Level 2-CE Prediction Report")
cat(" [Advisor]\n")
```

PID: `r data$PID`; Class of `r data$cohort`

```{r}
#| include: false
library(tidyverse)
library(scales)
observed_vars <- NULL
imputed_vars <- NULL

if(("COMAT_EM_imp" %in% names(data)) && data$COMAT_EM_imp) imputed_vars <- c("COMAT_EM", imputed_vars) else observed_vars <- c("COMAT_EM", observed_vars)
if(("COMAT_FM_imp" %in% names(data)) && data$COMAT_FM_imp) imputed_vars <- c("COMAT_FM", imputed_vars) else observed_vars <- c("COMAT_FM", observed_vars)
if(("COMAT_OPP_imp" %in% names(data)) && data$COMAT_OPP_imp) imputed_vars <- c("COMAT_OPP", imputed_vars) else observed_vars <- c("COMAT_OPP", observed_vars)
if(("COMAT_OBGYN_imp" %in% names(data)) && data$COMAT_OBGYN_imp) imputed_vars <- c("COMAT_OBGYN", imputed_vars) else observed_vars <- c("COMAT_OBGYN", observed_vars)
if(("COMAT_IM_imp" %in% names(data)) && data$COMAT_IM_imp) imputed_vars <- c("COMAT_IM", imputed_vars) else observed_vars <- c("COMAT_IM", observed_vars)
if(("COMAT_PEDS_imp" %in% names(data)) && data$COMAT_PEDS_imp) imputed_vars <- c("COMAT_PEDS", imputed_vars) else observed_vars <- c("COMAT_PEDS", observed_vars)
if(("COMAT_SURG_imp" %in% names(data)) && data$COMAT_SURG_imp) imputed_vars <- c("COMAT_SURG", imputed_vars) else observed_vars <- c("COMAT_SURG", observed_vars)
if(("COMSAE_att1_imp" %in% names(data)) && data$COMSAE_att1_imp) imputed_vars <- c("COMSAE_att1", imputed_vars) else observed_vars <- c("COMSAE_att1", observed_vars)
if(("Cumulative_Score_imp" %in% names(data)) && data$Cumulative_Score_imp) imputed_vars <- c("Cumulative_Score", imputed_vars) else observed_vars <- c("Cumulative_Score", observed_vars)
if(("COMLEX1_imp" %in% names(data)) && data$COMLEX1_imp) imputed_vars <- c("COMLEX1", imputed_vars) else observed_vars <- c("COMLEX1", observed_vars)

```

### Observed Student Data on Predictors of COMLEX-USA Level 2-CE

```{r}
#| echo: false
#| include: !expr ncol(data[observed_vars]) > 0

new_names <- c("COMLEX1 Performance" = "COMLEX1",
               "COMSAE_Attempt_1" = "COMSAE_att1")

observed_data <- as.data.frame(data[observed_vars], col.names = observed_vars) |> 
  dplyr::rename(any_of(new_names))
kableExtra::kable(observed_data, format = "latex", 
                  col.names = gsub("[_]", " ", names(observed_data)), 
                  row.names = FALSE,
                  align = "c") %>%
  kable_styling(font_size = 9, position = "left", 
                latex_options = c("HOLD_position", "scale_down")) |> 
  column_spec(1:ncol(observed_data),width="6.5em")
```

The table above shows observed measures of student performance available when we
ran this report (table omitted if no predictors were observed). Those may
include the first-attempt COMLEX-USA Level 1 exam result, cumulative score in
the COM program, first-attempt score on the COMSAE, and scores on various COMAT
exams.

```{r}
#| echo: false
#| results: "asis"
#| include: !expr advising

cat("### Expected Scores on Missing Predictors of COMLEX-USA Level 2-CE")
cat("\n")
```

```{r}
#| echo: false
#| include: !expr advising & ncol(data[imputed_vars]) > 0

scenarios <- as.data.frame(data[imputed_vars]) %>% 
  mutate(across(where(is.numeric), ~round2(.x, digits = 0)))
kableExtra::kable(scenarios, format = "latex", 
                  col.names = gsub("[_]", " ", names(scenarios)), 
                  row.names = FALSE,
                  align = "c") %>%
  kable_styling(font_size = 9, position = "left", 
                latex_options = c("HOLD_position", "scale_down"))
```

```{r}
#| echo: false
#| include: !expr advising

ExplainExpected <- paste("Our prediction model uses COM exam scores such as COMSAE, COMLEX-USA Level 1 and all COMATs but",
                         data$First_Name, 
                         "may not have taken all of them yet.", 
                         "To proceed despite missing values on some predictors,", 
                         "we used the observed data in the prior section to",
                         "impute the approximate scores", 
                         data$First_Name, 
                         "may achieve on the missing measures.", 
                         "Any resulting expected scores are shown in the table",
                         "above (table omitted if all predictors were observed).")
```

`r if(advising) ExplainExpected`

### Predictions About COMLEX-USA Level 2-CE Exam Performance

Using observed student performance, expected performance on missing exams, and
past COM cohort performance, we can estimate `r data$First_Name`'s performance
on the COMLEX-USA Level 2-CE exam. We note that prediction accuracy increases as
more student data is observed.

```{r}
#| echo: false
interval_vars <- c("COMLEX2_Pred", "Lower_95CI", "Upper_95CI")
interval <- data.frame(data[,interval_vars]) %>% 
  mutate(across(everything(), ~round2(.x, digits = 0)))
kableExtra::kable(interval, format = "latex", 
                  col.names = c("Point Prediction", "Interval Lower Bound", 
                                "Interval Upper Bound"), 
                  row.names = FALSE,
                  align = "c") %>%
  add_header_above(header = c("Predicted Score and 95% Prediction Interval" = 3)) %>%
  kable_styling(font_size = 9, position = "left", 
                latex_options = "HOLD_position")
```

The table above shows that `r data$First_Name` will most likely achieve
a score of about `r round2(data$COMLEX2_Pred, digits = 0)` on the
COMLEX-USA Level 2-CE exam, with a 95% chance of scoring between
`r round2(data$Lower_95CI, digits = 0)` and
`r round2(data$Upper_95CI, digits = 0)`.

```{r}
#| echo: false
#| warning: true

threshold_vars <- c("prob_400", "prob_450", "prob_500", "prob_550", "prob_600", 
                    "prob_650")
predictions <- data.frame(data[,threshold_vars]*100) %>% 
  mutate(across(everything(), ~round2(.x, digits = 0)))

kableExtra::kable(predictions, format = "latex", escape = FALSE, 
                  col.names = c("Score $\\geq 400$", "Score $\\geq 450$", 
                                "Score $\\geq 500$", "Score $\\geq 550$", 
                                "Score $\\geq 600$", "Score $\\geq 650$"), 
                  row.names =FALSE,
                  align = "c") %>%
  add_header_above(header = c("Percent Chance of Scoring Above Various Thresholds" = 6)) %>%
  kable_styling(font_size = 9, position = "left", 
                latex_options = "HOLD_position")%>%
  add_footnote("Values of 100% are artifacts of rounding percentages of 99.5 or higher to the nearest integer.")
```

Finally, the table above shows additional predictions about
`r data$First_Name`'s chances of scoring at or above several different
COMLEX-USA Level 2-CE score thresholds that may be relevant when applying to
various specialty programs. For example, `r data$First_Name` has a
`r predictions[1,"prob_400"]`% chance of passing (scoring above 400) and
a `r predictions[1,"prob_500"]`% chance of scoring above 500 on the
exam.

```{r}
#| echo: false
#| results: "asis"

cat(":::{.callout-tip}")
cat("\n")
cat("## Important Notice")
cat("\n")
cat("These predictions provide useful insight into *likely* exam outcomes but \n")
cat("*they are statistical estimates, not guarantees*.\n")
cat("The predictions assume that students will continue to exert similar levels of \n ")
cat("effort in studying and preparing for exams as they have in the past. \n")
cat("Even with complete data about all predictors, it is impossible to predict with 
    absolute certainty how individual students will perform.\n")
cat("The only way to get a definitive outcome is to take the COMLEX-USA Level 2-CE exam.\n")
cat(":::")
```
