---
title: "COMLEX2-CE Student Reports"
author:
  - name: Sam Manski
    email: manskisa@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
  - name: Anh Do
    email: doanh@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
  - name: Steven J. Pierce
    orcid: 0000-0002-0679-3019
    email: pierces1@msu.edu
    affiliations: 
      - name: Michigan State University, Center for Statistical Training and Consulting
execute-dir: file
params: 
  LogFile: "COMLEX2_Student_Reports_Draft.pdf"
format: 
  pdf:
    output-file: "COMLEX2_Student_Reports_Draft.pdf"
    output-ext: "pdf"
    documentclass: scrartcl
    papersize: letter
    geometry: 
      - top=0.75in
      - bottom=0.75in
      - left=0.75in
      - right=0.75in
    number-sections: false
    toc: true
    toc-depth: 2
    colorlinks: true
    template-partials: 
      - ../title.tex
    include-in-header:
      - file: ../compact-title.tex
      - text: |
          \usepackage{fancyhdr}
          \usepackage[noblocks]{authblk}
          \renewcommand*{\Authsep}{, }
          \renewcommand*{\Authand}{, }
          \renewcommand*{\Authands}{, }
          \renewcommand\Affilfont{\small}
          \usepackage[yyyymmdd,hhmmss]{datetime}
          \usepackage{lastpage}        %% For pageref command
          \usepackage{fontspec}
          \defaultfontfeatures{Ligatures=TeX}
          \usepackage[font={small}, margin=1cm, skip=2pt]{caption}
          \usepackage{placeins}        %% For FloatBarrier command
          \usepackage{booktabs}        %% Used by kableExtra
          \usepackage{longtable}       %% Used by kableExtra
          \usepackage{multirow}        %% Used by kableExtra
          \usepackage{wrapfig}         %% Used by kableExtra
          \usepackage{colortbl}        %% Used by kableExtra
          \usepackage{pdflscape}       %% Used by kableExtra
          \usepackage{tabu}            %% Used by kableExtra
          \usepackage[normalem]{ulem}  %% Used by kableExtra
          \usepackage{makecell}        %% Used by kableExtra
          \usepackage{xcolor}          %% Used by kableExtra
          \usepackage{dcolumn}         %% Used by kableExtra
          \usepackage{titling}      
          \usepackage{threeparttable}  %% Used by kableExtra
          \usepackage{threeparttablex} %% Used by kableExtra
          \usepackage{verbatim}        %% For comment command
          \usepackage{here}
          \pretitle{\begin{center}\LARGE\bfseries}
          \posttitle{\end{center}}
          \pagestyle{fancy}
          \lhead{COMLEX-USA Level 2-CE Score Prediction Model}
          \rhead{\today\ \currenttime}
          \cfoot{ }
          \fancyfoot[R]{\thepage\ of \pageref*{LastPage}}
          \renewcommand{\headrulewidth}{0.4pt}
          \renewcommand{\footrulewidth}{0.4pt}
          \fancypagestyle{plain}{\pagestyle{fancy}}
          \newcommand*\tocentryformat[1]{{\sffamily#1}}  %% Fix TOC font style
          \RedeclareSectionCommands                      %% Fix TOC font style
            [
              tocentryformat=\tocentryformat,
              tocpagenumberformat=\tocentryformat
            ]
            {section,subsection,subsubsection,paragraph,subparagraph}
execute: 
  eval: true 
  echo: false 
  output: true 
  message: true 
  warning: true 
  error: true 
  include: true 
knitr:
  opts_chunk: 
    message: true  
    cfsize: "scriptsize" 
---

\lfoot{\texttt{\small \detokenize{`r params$LogFile`}}}

```{r}
#| label: global-options

# Create a custom chunk hook/option for controlling font size in chunk & output.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$cfsize != "normalsize", 
         paste0("\n \\", options$cfsize,"\n\n", x, "\n\n \\normalsize"), 
         x)
  })
```

{{< pagebreak >}}

# Data Source

```{r}
#| label: load-packages-log
#| include: false

library(here)         # for here()
library(quarto)       # for quarto_render(), quarto_version()
library(lubridate)    # for today()
library(tidyr)        # for pivot_longer()
library(rmarkdown)    # for pandoc_version()
library(knitr)        # for kable()
opts_knit$set(eval.after = 'fig.cap') 
opts_chunk$set(warning = FALSE, cache.extra = rand_seed)
library(COMReports)   # for git_report()
library(webexercises) # for round2()
library(kableExtra)
library(tidyverse)        # for dplyr, ggplot2.
```

```{r}
#| label: define-render-child

render_child <- function(data, i, advising) {
  res = knitr::knit_child(
   text = xfun::read_utf8(here(paste0("scripts/Prediction/Advising_Template_Predict_COMLEX2.qmd"))),
   envir = rlang::env(data = data, i = i, advising = advising),
   quiet = TRUE
  )
  cat(res, sep = '\n')
  cat("\n")
}
```

```{r}
#| label: load-data
#| echo: fenced 
#| include: true

# Data file created by "scripts/Prediction/Predict_COMLEX2_Score.qmd"
load(file=here::here("data/Predict_COMLEX2_Score.RData"))

# Data file created by "scripts/Prediction/Predict_COMLEX2_Score.qmd"
load(file=here::here("data/COMBANK_All.RData"))

# Show the data file modification date and time. 
file.mtime(here::here("data/Predict_COMLEX2_Score.RData"))
```

{{< pagebreak >}}

# Prediction Variability

```{r}
#| include: false
median_student_name <- data.frame("Median", "Student", "APID000002", "2025", "Pass")
names(median_student_name) <- c("First_Name", "Last_Name", "PID", "cohort", "COMLEX1")
median_student_scores <- df_current %>% summarise(across(where(is.numeric)|where(is.logical),median))
median_student <- cbind(median_student_name, median_student_scores)
df_for_reports <- rbind(median_student, df_current)
median_student_est_sd = (median_student$Upper_95CI - median_student$COMLEX2_Pred)/2

historic_COMBANK <- COMBANK_All %>% filter(cohort != "2025")
historic_COMBANK$cohort <- droplevels(historic_COMBANK$cohort)

COMLEX2_summary <- data.frame(Median = median(historic_COMBANK$COMLEX2, na.rm = TRUE),
                              Mean = mean(historic_COMBANK$COMLEX2, na.rm = TRUE),
                              SD = sd(historic_COMBANK$COMLEX2, na.rm = TRUE),
                              Min = min(historic_COMBANK$COMLEX2, na.rm = TRUE),
                              Max = max(historic_COMBANK$COMLEX2, na.rm = TRUE),
                              Lower = quantile(historic_COMBANK$COMLEX2, probs = c(0.025), na.rm = TRUE),
Upper = quantile(historic_COMBANK$COMLEX2, probs = c(0.975), na.rm = TRUE), row.names = NULL)
```

This page discusses the variability present in COMLEX-USA Level 2-CE student
scores and predictions. When attempting to predict COMLEX-USA Level 2-CE scores
for individual students, we have two main sources of variability:

1. Natural variability in COMLEX-USA Level 2-CE scores between students. 
2. Uncertainty in the strength of relationships between predictor variables and 
   COMLEX-USA Level 2-CE scores due to sampling error.

## Natural Variability in Scores
First, scores on COMLEX-USA Level 2-CE are historically variable within the COM 
program at MSU. @tbl-historic-summary shows that for the 
`r min(as.numeric(levels(historic_COMBANK$cohort)))` to 
`r max(as.numeric(levels(historic_COMBANK$cohort)))` cohorts, the observed 
scores ranged from `r COMLEX2_summary$Min` to `r COMLEX2_summary$Max` with a 
mean of `r round(COMLEX2_summary$Mean, digits = 2)` and a standard deviation of 
`r round(COMLEX2_summary$SD, digits = 2)`. If we ignore potential predictors, 
we could naively expect a randomly selected student to obtain a score within the 
95% empirical interval shown in @tbl-historic-summary, which is about 
`r round(COMLEX2_summary$Upper - COMLEX2_summary$Lower, digits = 0)` points 
wide. That extremely wide interval motivates our interest in using valid 
predictors of the test scores. A good model should give us a narrower (more 
precise) prediction interval that reduces the uncertainty about the student's 
expected score. 

```{r}
#| echo: false
#| label: tbl-historic-summary
#| tbl-cap: Summary statistics for historic COMLEX-USA Level 2-CE scores at MSU COM (Cohorts 2021-2024)

kable(COMLEX2_summary, digits = 2) %>% 
  add_header_above(c(" " = 5, "95% Empirical Interval" = 2))
```

Our model estimates how much residual variability remains between students after
accounting for the effects of the predictors (COMSAE, COMLEX-USA Level 1, and
COMAT scores). It is exceedingly rare for a model to explain all of the 
natural variability observed in the outcome. *Thus, even two students that have 
identical values for all predictors will likely score differently on COMLEX-USA 
Level 2-CE*. Our predictions must reflect that residual variability to be 
realistic. 

## Sampling Error in Estimates of Predictor Effects
The second source of variability comes from the sampling error associated with 
estimating relationships between the model predictors (Cumulative score, COMSAE 
first attempt, COMLEX-USA Level 1 first attempt, and COMAT subject exams) and 
COMLEX-USA Level 2-CE scores in a sample of students rather than from the entire 
population of interest. The uncertainty regarding the exact values of these
regression coefficients in the population also contributes some variability to
predicted COMLEX-USA Level 2-CE scores.

## How Prediction Intervals Are Obtained
For each student, our model simulates 6,000 possible COMLEX-USA Level 2-CE exam
results, accounting for both variability in the impact of predictors and 
residual variability between students. We then summarize these results into a
point prediction (the median of simulated scores) and a 95% prediction interval
(the 2.5th and 97.5th percentiles of the simulated scores). There is a 95% 
chance that the individual student will get a score within this interval. 
Because the distributions of predicted values for a given student are 
approximately normal, the bounds of the prediction interval should be about
$\pm 1.96$ standard deviations away from the point prediction score.

## Comparing Prediction Intervals to Historic Distribution of Exam Scores

```{r}
#| label: fig-density
#| fig-height: 2
#| fig-width: 7
#| echo: false
#| fig-cap: Historic COMLEX-USA Level 2-CE distribution and 95% interval contrasted with predicted distribution and 95% interval for a median student.
ggplot(data = historic_COMBANK) + 
  geom_density(aes(x = COMLEX2, col = "Historic_Distribution")) +
  geom_function(fun = dnorm, 
                args = list(mean = median_student$COMLEX2_Pred, 
                            sd = median_student_est_sd), 
                aes(col = "Median_Student_Predicted"))+
  geom_vline(aes(xintercept = median_student$Lower_95CI, 
                 col = "Median_Student_Predicted"), 
             linetype = "dotted")+ 
  geom_vline(aes(xintercept = median_student$Upper_95CI, 
                 col = "Median_Student_Predicted"), 
             linetype = "dotted")+
  geom_vline(aes(xintercept = COMLEX2_summary$Min, col = "Historic_Min_Max"))+ 
  geom_vline(aes(xintercept = COMLEX2_summary$Max, col = "Historic_Min_Max"))+
  geom_vline(aes(xintercept = COMLEX2_summary$Lower, 
                 col = "Historic_Distribution"), 
             linetype = "dotted")+ 
  geom_vline(aes(xintercept = COMLEX2_summary$Upper, 
                 col = "Historic_Distribution"), 
             linetype = "dotted")+
  scale_color_manual(name = " ", 
                     values = c(Historic_Distribution = "black", 
                                Median_Student_Predicted = "blue", 
                                Historic_Min_Max = "orange"))
```

We can contrast the predicted interval for a given student with the historic
distribution of COMLEX-USA Level 2-CE scores to show that though our prediction
intervals may still encompass a large range of scores, they are much narrower 
than the historic distribution. @fig-density shows the historic distribution of 
COMLEX-USA Level 2-CE scores as the black solid line with the vertical dotted 
black lines representing the empirical 95% interval. We then overlay a predicted 
distribution for a median student in blue with vertical dotted blue lines 
representing the 95% prediction interval. Finally, we include the historic 
minimum and maximum score observed as vertical orange lines to show the full 
range of observed COMLEX-USA Level 2-CE scores. 

::: {.callout-tip}
The prediction interval is narrower than the historic distribution precisely 
because the model includes useful and informative predictors. However, it does 
not collapse down to a single point because the available predictors do not 
explain all of the natural variability in student scores and there is sampling 
error in the estimated effects of the predictors. 
:::

If our interest lay solely in the expected mean score for students with a
particular combination of values on the predictors, we could produce a 95%
confidence interval for the *mean* student score that would be narrower than the
prediction interval for an individual student's predicted score. However, here
we aim to make predictions regarding individuals, so we must include the
residual variance when setting the interval bounds. That is why the prediction
intervals for a single student are wider than the confidence interval for the mean.

Finally, we note that as more data is collected, both for individual student 
predictors and COMLEX-USA Level 2-CE results for entire cohorts, the width of 
our prediction intervals should decrease because sampling error is a function of 
the sample size used to estimate the model. 

{{< pagebreak >}}

# Individual Reports

```{r}
#| results: "asis"
#| include: true



#for(i in 1:1) # For testing purposes
for(i in 1:length(df_for_reports$PID)) # For complete report
{
  render_child(data = df_for_reports[i,], i = i, advising = TRUE)
  render_child(data = df_for_reports[i,], i = i, advising = FALSE)
}
```
